---
title: "Regression Analysis"
output:   
  html_document:
    toc: true
    toc_float: true
---

```{r,include=FALSE,message=FALSE,echo=FALSE echo=FALSE, message=FALSE, r,include=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(plotly)
library(highcharter)
library(rworldmap)
library(gganimate)
library(transformr)
library(corrplot)
library(leaps)
library(kableExtra)
library(dplyr)
knitr::opts_chunk$set(
 echo = FALSE,
 fig.width = 7, 
 fig.height = 5,
 fig.asp = 0.6,
 out.width = "60%")
theme_set(theme_bw() + 
          theme(legend.position = "bottom",
                legend.title = element_blank(),
                plot.title = element_text(hjust = 0.5, size = 15),
                plot.subtitle = element_text(hjust = 0.5, size = 12)))
```

### Data Import and Cleaning

#### Death dataset.
```{r}
deaths = read_csv("./data/final_hiv_deaths.csv") %>% 
  janitor::clean_names() %>% 
  select("measure_name",
    "location_name", starts_with("sex"), starts_with("age"), "year", "val", "upper", "lower") %>% 
  rename(country_name = location_name)
  
deaths %>% 
  summarize(n_countries = n_distinct(country_name),
            n_years = n_distinct(year))

```

The death dataset comes from:
It contains HIV death count data from 204 distinct countries, in the 30 years between 1990 and 2019.

#### GDP and population datasets.

```{r}
gdp = read_csv("./data/final_gdp_per_capita.csv", skip = 4, col_names = TRUE) %>% 
  pivot_longer(cols = "1960":"2021",
               names_to = "year",
               values_to = "GDP per capita") 

population = read_csv("./data/final_population_total.csv", skip = 4, col_names = TRUE) %>% 
  pivot_longer(cols = "1960":"2021",
               names_to = "year",
               values_to = "Population")

gdp_population = inner_join(gdp, population, by = c("Country Name", "year")) %>% 
  janitor::clean_names() %>% 
  select("country_name", "year", "gdp_per_capita", "population") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year>=1990, year<=2019) 

aa = gdp_population %>% 
  select("year","population") %>% 
  arrange(year) %>% 
  group_by(year) %>% 
  summarize(pop = sum(population,na.rm=T))


```

The GDP and population datasets both come from:
After combining, the gdp_population dataset contains data from 266 distinct countries, in the 62 years between 1960 and 2021.

#### Merging datasets to produce the "maindata" dataset for analysis purpose.

```{r}
maindata = left_join(deaths, gdp_population, by = c("year","country_name")) 

  head(maindata,100)
```


population by year
```{r}
pop_plot = gdp_population %>% 
ggplot(aes(x = population))  
```

#### Description
### Global trend of HIV mortality per 100k population overtime (1990 to 2019)
```{r}
death_year = maindata %>% 
  group_by(year) %>% 
  summarize(HIV_death = sum(val),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
         motality_per_100k =     HIV_death/(pop_year/100000)) 


view(death_year)


death_year %>% 
  ggplot(aes(x = year, y = motality_per_100k, color = year)) + geom_point()
```


### Global trend of HIV mortality per 100k population by gender overtime (1990 to 2019)
```{r}
death_gender = maindata %>% 
  group_by(year, sex_name) %>% 
  summarize(HIV_death = sum(val),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
         motality_per_100k =     HIV_death/(pop_year/100000))

death_gender %>% 
  ggplot(aes(x = year, y = motality_per_100k, color = sex_name)) + geom_point()
```

### Global trend of HIV mortality per 100k population by age overtime (1990 to 2019)
```{r}

```

rate_year = left_join(death_year, gdp_population, by = c("year")) %>%
  group_by(year) %>% 
  mutate(
    HIV_mortality = HIV_death/(population/100000))
  
  
  
pop_year = gdp_population %>%
  mutate(population = as.numeric(population)) %>% group_by(year) %>% 
  summarize(year_pop = sum(population))

  
```


### Sum of HIV death by country and year
```{r}
death_sum = maindata %>% 
  group_by(country_name,year) %>% 
  summarize(HIV_death = sum(val)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death)
  )

 
  
```

#### HIV/AIDS deaths per 100,000 population
```{r}
mortality = left_join(death_sum, gdp_population, by =c("year","country_name")) %>% 
  filter(year<=2021,year>=1990) %>% 
  mutate(
    HIV_mortality = HIV_death/(population/100000))
```


  

head(mortality,100)

maindata %>% mutate(
    HIV_death = as.integer(val)
  ) 
```

